Evolution of Processes in Delivering Reliable Software in the European Scientific Arena

### 2018-08-23 paper-ieee-trans.tex
Caption text elaborated with an explanation of zero-value coverage for products in INDIGO-1

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{images/fig2.png}
\caption{Comparison of the unit testing coverage values for the {\sl INDIGO--DataCloud}
software stack over the {\sl INDIGO--1} and {\sl INDIGO--2} major releases.}
\label{fig:fig_unittest}
\end{figure*}

@@

\item \textit{Documentation}.
The SQA criteria set the path for the adoption of the developed software by
defining the documentation content, according to the target audience. This requirement also
promotes the automation, both in terms of documentation creation and service deployment. The
\textit{documentation is treated as code}, using a markup language, automatically rendered and
uploaded to online repositories\footnote{\url{https://www.gitbook.com/@indigo-dc}}.
Thus documentation is portable and
human--reviewed, using the same workflow as the code does, validating the changes before being
updated into the production repository.

\item \textit{Automated Deployment}.
To lower the barriers of software adoption, the SQA criteria require the automated
deployment of the products delivered as part of the catalogue. Automation in this context is
tackled using configuration management tools. These tools allow the adoption of
\textit{Infrastructure as Code} (IaC) practice, managing the component's deployment through declarative
definitions. The definition files appear as an additional source of documentation
-- self--documenting code -- as they sequentially guide the component's deployment process on multiple
platforms. {\sl INDIGO--DataCloud} project contributed to open--source IaC tools such
as Ansible
and Puppet\footnote{\url{https://forge.puppet.com/indigodc}}. A representative example of
such contributions are the \textit{50 roles} developed from scratch and currently hosted in
the Ansible Galaxy portal\footnote{\url{https://galaxy.ansible.com/indigo-dc/}}. Fig.~\ref{fig:fig_confman}
shows the number of products that offer an automated means for deployment.
It shows an increase in the adoption of such tools leading to the {\sl INDIGO--1} release
(light cream points and trend line), as well as to the {\sl INDIGO--2} release
(dark blue points and trend line). The rate of adoption is lower during the weeks
before the second release because a significant fraction of the products had already adopted it
previously to the first release.
\end{itemize}

@@

\textit{Development metrics} are obtained programmatically from several sources, namely GitHub
API\footnote{\url{https://developer.github.com/v3/}.}
and the Jenkins\footnote{\url{https://jenkins.io/}} service, and graphically displayed as GitHub
pages using the GrimoireLab framework\footnote{\url{http://grimoirelab.github.io/}}.
Per--component weekly reports, including the SQA requirement
fulfilment, are issued and individually discussed through the different communication channels
with the development teams. Issue tracking metrics and KPIs are an essential piece of information
of both feature addition and defect solving, useful to detect and fix misbehaviours while in the
development phase.

\textit{Release metric} sources are the online repository servers, namely the
Linux package\footnote{\url{https://jenkins.indigo-datacloud.eu/}}
and DockerHub\footnote{\url{https://hub.docker.com/u/indigodatacloud}} repositories. Jenkins
server also contains valuable release data as it is the service where the packages
and containers are being built before being uploaded to the online repositories. Release
KPIs primarily focus on the frequency and efficiency -- mainly rollbacks -- statistics.

@@

The practical implementation of the CI scenario leverage from a tightly integrated open
source tools such as GitHub\footnote{\url{https://github.com/}}, as the online source code repository,

@@

>>>>

\begin{figure*}[ht]
\centering
\includegraphics[width=\textwidth]{images/fig2.png}
\caption{Comparison of the unit testing coverage values for the {\sl INDIGO--DataCloud}
software stack over the {\sl INDIGO--1} and {\sl INDIGO--2} major releases. Products that
were exclusively released during {\sl INDIGO--2} do not show data for the first period of
{\sl INDIGO--1}.}
\label{fig:fig_unittest}
\end{figure*}

@@

\item \textit{Documentation}.
The SQA criteria set the path for the adoption of the developed software by
defining the documentation content, according to the target audience. This requirement also
promotes the automation, both in terms of documentation creation and service deployment. The
\textit{documentation is treated as code}, using a markup language, automatically rendered and
uploaded to online repositories \cite{foot1}.
Thus documentation is portable and
human--reviewed, using the same workflow as the code does, validating the changes before being
updated into the production repository.

\item \textit{Automated Deployment}.
To lower the barriers of software adoption, the SQA criteria require the automated
deployment of the products delivered as part of the catalogue. Automation in this context is
tackled using configuration management tools. These tools allow the adoption of
\textit{Infrastructure as Code} (IaC) practice, managing the component's deployment through declarative
definitions. The definition files appear as an additional source of documentation
-- self--documenting code -- as they sequentially guide the component's deployment process on multiple
platforms. {\sl INDIGO--DataCloud} project contributed to open--source IaC tools such
as Ansible
and Puppet \cite{foot2}. A representative example of
such contributions are the \textit{50 roles} developed from scratch and currently hosted in
the Ansible Galaxy portal \cite{foot3}. Fig.~\ref{fig:fig_confman}
shows the number of products that offer an automated means for deployment.
It shows an increase in the adoption of such tools leading to the {\sl INDIGO--1} release
(light cream points and trend line), as well as to the {\sl INDIGO--2} release
(dark blue points and trend line). The rate of adoption is lower during the weeks
before the second release because a significant fraction of the products had already adopted it
previously to the first release.
\end{itemize}

@@

\textit{Development metrics} are obtained programmatically from several sources, namely GitHub
API \cite{foot4} and the Jenkins \cite{foot5} service, and graphically displayed as GitHub
pages using the GrimoireLab framework \cite{foot6}.
Per--component weekly reports, including the SQA requirement
fulfilment, are issued and individually discussed through the different communication channels
with the development teams. Issue tracking metrics and KPIs are an essential piece of information
of both feature addition and defect solving, useful to detect and fix misbehaviours while in the
development phase.

\textit{Release metric} sources are the online repository servers, namely the
Linux package \cite{foot7} and DockerHub \cite{foot8} repositories. Jenkins
server also contains valuable release data as it is the service where the packages
and containers are being built before being uploaded to the online repositories. Release
KPIs primarily focus on the frequency and efficiency -- mainly rollbacks -- statistics.

@@

The practical implementation of the CI scenario leverage from a tightly integrated open
source tools such as GitHub \cite{foot9}, as the online source code repository,

@@

\bibitem{foot1}
indigo-dc, ``Sign in to GitBook - GitBook,''
2018.
[Online]. Available: \url{https://www.gitbook.com/@indigo-dc}.
[Accessed 30 Aug. 2018].

\bibitem{foot2}
puppetforge, ``Modules by INDIGO Datacloud - Puppet Forge,''
2018.
[Online]. Available: \url{https://forge.puppet.com/indigodc}.
[Accessed 30 Aug. 2018].

\bibitem{foot3}
GALAXY, ``Ansible Galaxy,''
2018.
[Online]. Available: \url{https://galaxy.ansible.com/indigo-dc/}
[Accessed 30 Aug. 2018].

\bibitem{foot4}
GitHub Developer, ``GitHub API v3 | GitHub Developer Guide,''
2018.
[Online]. Available: \url{https://developer.github.com/v3/}
[Accessed 30 Aug. 2018].

\bibitem{foot5}
Jenkins, ``Jenkins,''
2018.
[Online]. Available: \url{https://jenkins.io/}
[Accessed 30 Aug. 2018].

\bibitem{foot6}
GRIMOIRELAB, ``GrimoireLab - Software Development and Community Analytics Platform,''
2017.
[Online]. Available: \url{http://grimoirelab.github.io/}
[Accessed 30 Aug. 2018].

\bibitem{foot7}
Jenkins Indigo-dc, ``Jenkins - Indigo-DataCloud,''
2018.
[Online]. Available: \url{https://jenkins.indigo-datacloud.eu/}
[Accessed 30 Aug. 2018].

\bibitem{foot8}
Indigo-dc, ``indigodatacloud - Docker Hub,''
2018.
[Online]. Available: \url{https://hub.docker.com/u/indigodatacloud}
[Accessed 30 Aug. 2018].

\bibitem{foot9}
GitHub, ``Learn Git and GitHub without any code!,''
2018.
[Online]. Available: \url{https://github.com/}
[Accessed 30 Aug. 2018].
